---
title: "Project_3"
author: "Koohyar P, Anthony C, Victor T, & James N"
date: "2024-03-16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(rvest)
library(stringr)
library(tidyr)
library(XML)
library(dplyr)
library(purrr)
library(stringr)
library(kableExtra)
library(reactable)
library(tidyverse)
```

### Data Import
https://ballotpedia.org/Super_Tuesday_primaries,_2024

```{r import}
url <- "https://ballotpedia.org/Super_Tuesday_primaries,_2024"
webpage <- read_html(url)
st_table <- html_nodes(webpage, '.portal-section')
table_names <- c("Alabama", "Alaska", "American Somoa", "Arkansas", "California", "Colorado", "Iowa", "Maine", "Massachusetts", "Minnesota", "North_Carolina", "Oklahoma", "Tennessee", "Texas", "Utah", "Vermont", "Virginia")
table_frames <- list()
for (i in seq_along(st_table)) {
  table_data <- html_table(st_table[[i]])
  table_name <- table_names[i]
  table_frames[[table_name]] <- table_data
}
# Sample Data Frame
table_frames$Alabama

```

### Remove Extraneous Rows

```{r removing rows with 99% reporting or Source}
for (i in seq_along(table_frames)){
  # Filter the rows that contain the character string "% reporting"
  table_frames[[i]] <- table_frames[[i]][!grepl("% reporting", table_frames[[i]]$X2), ]
}
# Loop through each data frame removing the row that contains "Source" in column X2
for (i in seq_along(table_frames)){
  # Filter the rows that contain the character string "Source"
  table_frames[[i]] <- table_frames[[i]][!grepl("Source", table_frames[[i]]$X2), ]
}
# for (i in seq_along(table_frames)) {
#  print(table_frames[[i]])
#}
table_frames[[1]]
```

### Adding Columns for State and Party affiliation

```{r adding columns for state and party affilation, echo=FALSE}
add_columns <- function(df){
  df |> mutate(X6 = if_else(X2 == "Candidate", "state", NA),
               X7 = if_else(X2 == "Candidate", "party", NA))
}
table_frames <- map(table_frames, add_columns)

table_frames[[1]]
```

### Extracting of Data Frames from List for Super Tuesday States
Selecting the proper columns, dropping the unnecessary ones.  This also places "Democrat into the Party Column for ALL candidates, which will fixed in a later section to include Republican in the correct candidate.  Pulls the individual state data frames out of the list to be worked on individually.

```{r super tuesday states, echo=FALSE}
alabama <- table_frames[[1]] |> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "alabama", X6))|> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
arkansas <- table_frames[[4]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "arkansas", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
california <- table_frames[[5]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "california", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
colorado <- table_frames[[6]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "colorado", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
maine <- table_frames[[8]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "maine", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
massachusets <- table_frames[[9]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "massachusets", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
minnesota <- table_frames[[10]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "minnesota", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
north_carolina <- table_frames[[11]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "north_carolina", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
oklahoma <- table_frames[[12]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "oklahoma", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
tennessee <- table_frames[[13]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "tennnessee", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
texas<- table_frames[[14]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "texas", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
utah <- table_frames[[15]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "utah", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
vermont <- table_frames[[16]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "vermont", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))
virginia <- table_frames[[17]]|> select(!X1&!X5) |> mutate(X6 = ifelse(is.na(X6), "virginia", X6)) |> mutate(X7 = ifelse(is.na(X7), "Democrat", X7))

alabama
```

### Substituting Republican in the rows
This uses a filter that identifies the second occurrence in the column X2 that separates republican candidates from the democratic candidates, and then replaces "Democrat" with "Republican".  This is done for each individual data frame for the Super Tuesday states. 
```{r repubilcan candidates, echo=FALSE}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- alabama$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  alabama <- alabama %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- arkansas$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  arkansas <- arkansas %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- california$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  california <- california %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- colorado$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  colorado <- colorado %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- maine$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  maine <- maine %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- massachusets$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  massachusets <- massachusets %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- minnesota$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  minnesota <- minnesota %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- north_carolina$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  north_carolina <- north_carolina %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- oklahoma$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  oklahoma <- oklahoma %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- tennessee$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  tennessee <- tennessee %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- texas$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  texas <- texas %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- utah$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  utah <- utah %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- vermont$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  vermont <- vermont %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
# Find the index of the second occurrence of "Candidate" in column X2
soi <- virginia$X2 %>%
  grep("Candidate", ., fixed = TRUE) %>%
  .[2]
# Replace values in column X7 after the second occurrence of "Candidate"
if (!is.na(soi)) {
  virginia <- virginia %>%
    mutate(X7 = ifelse(row_number() > soi, "Republican", X7))
}
alabama
```

### Removing Candidate Rows
In this section, we remove the "candidate" row from each of the data frames and rename the column variable names from X to what they actually represent.  We also address the accidental import of California candidates not running for the national office of president because they were placed in the same table on the website.  Those candidates had their party affiliation and percent_vote values changes to NA.
```{r removing candidate rows by filter, echo=FALSE, warning=FALSE}
alabama <- alabama |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
arkansas <-  arkansas |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
california <-  california |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
# California Cleanup. Five candidates brought over that were not running for president.
california[18:23, "party"] <- "NA"
california[18:23, "percent_vote"] <- "NA"
colorado <-  colorado |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
maine <-  maine |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
massachusets <-  massachusets |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
minnesota <-  minnesota |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
north_carolina <-  north_carolina |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
oklahoma <-  oklahoma |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
tennessee <- tennessee |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
texas <-  texas |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
utah <-  utah|> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
vermont <-  vermont |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
virginia <- virginia |> filter(!X2=="Candidate") |> rename(candidate=X2, percent_vote=X3, num_vote=X4, state=X6, party=X7)
california |> kbl() |> kable_classic_2(full_width=F, font_size=12)

```

### Combining the Entire Data Frame
In this section we combine all the Super Tuesday state data frames into one tidy data frame containing five (5) variables and 205 rows.  Once in, we will remove all the commas in the vote counts so we can use the data as numeric.
```{r Combine}
# The Combine
super_tuesday_combine <- bind_rows(alabama, arkansas, california, colorado, maine, massachusets, minnesota, north_carolina, oklahoma, tennessee, texas, utah, vermont, virginia)

# Remove the commas
super_tuesday_combine$num_vote <- gsub(",", "", super_tuesday_combine$num_vote)
super_tuesday_combine
```
### Total Votes Cast by Party Affiliation and State

In this section we are going to identify the total number of votes cast by state, and then grouped by party affiliation.

```{r state and party votes, echo=FALSE}
# Treat the num_vote column as numeric
super_tuesday_combine$num_vote <- as.numeric(super_tuesday_combine$num_vote)


#Group by state total votes
state_total_2024 <- super_tuesday_combine |> group_by(state) |> summarise(state_total=sum(num_vote))

state_dem_total_2024 <- super_tuesday_combine |> group_by(state,party) |> filter(party=="Democrat") |>  summarise(state_party_total_d=sum(num_vote), .groups = "keep")

state_rep_total_2024 <- super_tuesday_combine |> group_by(state, party) |> filter(party=="Republican") |>  summarise(state_party_total_r=sum(num_vote), .groups = "keep")

state_data_2024a <- merge(state_total_2024, state_dem_total_2024, by="state")
state_data_2024b <- merge(state_data_2024a, state_rep_total_2024, by="state")

state_final_2024 <- state_data_2024b |> select(!party.x&!party.y)

state_final_2024
```

### Importing Next Dataset
In this section we are now going to import the state level data on regsitered voters and party affiliations.  This comes from the websitehttps://worldpopulationreview.com/state-rankings/registered-voters-by-state.  These figures are from October 2022, but should be accurate enough for our purposes of comparison with the voting electorate on Super Tuesday 2024.  This import was much simpler due to the table structured format of the website, so the transformation and clean up are going to be much simpler.  The initial import required us to move the first row to column names and then remove the fist row as an observation.

```{r partisan affilation, echo=FALSE}
url <- "https://worldpopulationreview.com/state-rankings/registered-voters-by-state"
webpage <- read_html(url)
vote <- webpage |> html_table(header = NA, fill = TRUE)
vote_2022 <- html_table(webpage, header = TRUE)[[2]]
#colnames(vote_2022) <- vote_2022[1,]
#vote_2022 <- vote_2022[-1,]
#vote_2022
```

### Tidying and Transforming
In this section we are going to clean up the data frame a bit.  We need to remove the commas and % in the values so we can do calculations later.  We need to adjust the column names to be more workable, and we will need to mutate the population numbers to be comparable with the values in the other data frame (these are in thousands).   
```{r tidy and transform, echo=FALSE}
# Column names
new_colnames <- c("state", "reg_voter_num", "perc_vote_pop")
colnames(vote_2022) <- new_colnames

# Remove the commas and convert to numeric
vote_2022$reg_voter_num <- gsub(",", "", vote_2022$reg_voter_num)
vote_2022$reg_voter_num <- as.numeric(vote_2022$reg_voter_num)

# Convert units from per 1000
vote_2022 <- vote_2022 |> mutate(reg_voter_num=reg_voter_num*1000)

# Remove % sign
vote_2022$perc_vote_pop <- gsub("%", "", vote_2022$perc_vote_pop)

# Convert percent
vote_2022 <- vote_2022 |> mutate(perc_vote_pop = as.numeric(perc_vote_pop)/100)

# Convert state to lower case
vote_2022$state <- tolower(vote_2022$state)

```

### Filtering the State Level Data to Only Super Tuesday States

```{r filter state partisan voter date}
super_states <- state_final_2024$state
vote_2022af <- vote_2022[vote_2022$state %in% super_states, ]

combine_data_frame <- merge(state_final_2024, vote_2022af, by="state")

combine_data_frame

```


*** Split the state form numbers in first column
state_split <- strsplit(vote_2022$State, "\\[")
vote_2022$State <- sapply(state_split,"[", 1)

*** Select the columns we want to work with
vote_2022a <- select(vote_2022, "State", "Total registrants", "Democratic Party", "Republican Party", "Independent", "Other parties")

*** Rename the columns
new_colnames <- c("state", "total_voters", "democratic_voters", "republican_voters","independent_voters", "others")
colnames(vote_2022a) <- new_colnames

*** Remove the commas
columns_commas <- c("total_voters", "democratic_voters", "republican_voters", "independent_voters", "others")
vote_2022a[columns_commas] <- lapply(vote_2022a[columns_commas], function(x) gsub(",", "", x))
vote_2022a

### Merging both Data sets

```{r}
## Merging Super_tuesday_combine with combine_data_frame
merge_combine<- merge(super_tuesday_combine,combine_data_frame, by="state",all.x = TRUE) %>%
  arrange((state), percent_vote)

merge_combine
```

### Analysis

## Calculate Average

To better visualize the data I will provide a table that display state by their Average Total
```{r}
merge_combine_1 <- mutate(merge_combine, "Average_total"= rowMeans( select(merge_combine,"num_vote", "state_party_total_r", "state_party_total_r", "reg_voter_num") ) )

reactable(select(merge_combine_1, "state", "Average_total"))
```

### Filter Votes by Party
This filter data party which are Democrat and Republican. W
```{r}
# Members filter 
Party_D <- merge_combine_1 %>% filter( party == "Democrat", na.omit(TRUE))
Party_R <- merge_combine_1 %>% filter(party == "Republican", na.omit(TRUE))
```

## Democratic vs Republoican vote by state
There are two plot below, one show Republican votes by State and the other show Democratic votes by state.
```{r}
## Democratic by state
D_plot <- ggplot(merge_combine_1, aes(y=merge_combine_1$state, fill = state)) + geom_bar(binwidth = 300) +
  ggtitle("Democratic Vote By State") +
  ylab("State Vote") +
  xlab("Democratic Party")

D_plot

## Republican by state
R_plot <- ggplot(merge_combine_1, aes(y=merge_combine_1$state, fill = state)) + geom_bar(binwidth = 300) +
  ggtitle("Republican Vote By State") +
  ylab("State Vote") +
  xlab("Republican Party")

R_plot
```

```{r}
myplot <- ggplot(merge_combine_1, aes(merge_combine_1$party, group = state)) +
          geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5) +
geom_bar(aes(y = ..prop.., fill = merge_combine_1$state, stat="count")) + 
          scale_y_continuous(labels=scales::percent) +
          ylab("Percentage of Voters by State") +
          xlab("Party")
  
myplot
```

```{r,warning=FALSE,echo=FALSE}
ggplot(merge_combine_1, aes(x =`num_vote`, y= `state`, fill = party)) + geom_boxplot()
```

### Conclusion